Drivers of sex differences in consumption

### 2017-05-12 sex-difference/sex-difference.Rmd
sex diff

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=4, fig.path='fig/', dev=c('png', 'pdf'),
                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE)
```

```{r load_data, cache=TRUE}
usage = read_tsv('../bene_usage_2011.tsv')
```

@@

```{r compute_consumption_by_age_and_sex}
bounds_group = function(x, upper_limits) {
  upper_limits = as.integer(upper_limits)
  levels = lapply(seq(from=1, to=length(upper_limits)), function(i) {
    if (i == 1) {
      sprintf('<= %i', upper_limits[i])
    } else {
      sprintf('> %i and <= %i', upper_limits[i - 1], upper_limits[i])
    }
  })
  top_level = sprintf('> %i', tail(upper_limits, 1))
  formula_list = mapply(function(str, lim) as.formula(sprintf('x <= %i ~ "%s"', lim, str)),
                        levels, upper_limits) %>%
    c(as.formula(sprintf('TRUE ~ "%s"', top_level)))

  levels = c(levels, top_level)

  do.call(case_when, formula_list) %>%
    factor(levels=levels, ordered=TRUE)
}

usage %>%
  filter(sex %in% c('male', 'female')) %>%
  mutate(age_group=bounds_group(age, c(65, 75, 85, 95))) %>%
  group_by(age_group, sex, bene_id) %>%
  summarize_at(vars(n_claims, n_days), sum) %>%
  summarize(mean_n_claims=mean(n_claims),
            mean_n_days=mean(n_days),
            n_bene=n()) %>%
  kable
```

@@

```{r drugs_by_sex, cache=TRUE}
bene_by_sex = usage %>%
  #sample_frac(0.01) %>%
  select(bene_id, sex) %>%
  distinct %>%
  count(sex) %>%
  rename(n_bene=n)

dbs = usage %>%
  #sample_frac(0.01) %>%
  filter(age >= 65, sex %in% c('male', 'female')) %>%
  group_by(antibiotic, sex, bene_id) %>%
  summarize_at(vars(n_claims, n_days), sum) %>%
  summarize(n_claims=sum(n_claims), n_days=sum(n_days)) %>%
  left_join(bene_by_sex, by='sex') %>%
  mutate(cpkp=n_claims*1000/n_bene, did=n_days*1000/(365*n_bene))

dbs %>%
  filter(sex=='female') %>%
  arrange(desc(cpkp)) %>%
  head(10) %>%
  kable

dbs %>%
  filter(sex=='male') %>% 
  arrange(desc(cpkp)) %>%
  head(10) %>%
  kable
```

```{r drug_by_sex_plot}
dbs %>%
  select(antibiotic, sex, cpkp) %>%
  spread(sex, cpkp) %>%
  ggplot(aes(x=male, y=female, label=antibiotic)) +
  geom_label() +
  geom_point() +
  geom_abline() +
  xlab('claims per 1,000 male beneficiaries') +
  ylab('claims per 1,000 female beneficiaries') +
  theme_minimal()
```

>>>>

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=4, fig.path='fig/', dev=c('png', 'pdf'),
                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE, autodep=TRUE)
```

```{r load_data}
year = 2011
bene = read_tsv(sprintf('../bene_%i.tsv', year)) %>%
  filter(between(age, 66, 96), hmo_months==0, sex %in% c('male', 'female')) %>%
  mutate(is_female=sex=='female', is_white=race=='white')

pde = read_tsv(sprintf('../pde_%i.tsv', year)) %>%
  filter(bene_id %in% bene$bene_id)

total_pde = pde %>% count(bene_id) %>% rename(n_claims=n)
```

@@

```{r consumption_by_age_and_sex}
bene %>%
  left_join(total_pde, by='bene_id') %>%
  replace_na(list(n_claims=0)) %>%
  group_by(sex, age) %>%
  summarize(mean_n_claims=mean(n_claims)) %>%
  ungroup() %>%
  ggplot(aes(x=age, y=mean_n_claims, color=sex)) +
  geom_point()
```

@@

```{r drugs_by_sex}
usage_by_sex = function(drug) {
  bene %>%
    left_join(pde %>% filter(antibiotic==drug) %>% count(bene_id) %>% rename(n_claims=n), by='bene_id') %>%
    replace_na(list(n_claims=0)) %>%
    group_by(sex) %>%
    summarize(mean_n_claims=mean(n_claims)) %>%
    ungroup() %>%
    spread(sex, mean_n_claims) %>%
    mutate(drug=drug)
}

dbs = lapply(unique(pde$antibiotic), usage_by_sex) %>% bind_rows
```

```{r}
denom = bene %>% count(sex) %>% rename(n_denom=n)
ubs = pde %>%
  select(bene_id, antibiotic) %>%
  distinct() %>%
  left_join(select(bene, bene_id, sex), by='bene_id') %>%
  count(sex, antibiotic) %>%
  ungroup() %>%
  rename(n_user=n) %>%
  left_join(denom, by='sex') %>%
  mutate(frac_using=n_user/n_denom) %>%
  select(antibiotic, sex, frac_using) %>%
  spread(sex, frac_using)
```

```{r drug_by_sex_plot}
ubs %>%
  select(antibiotic, sex, cpkp) %>%
  spread(sex, cpkp) %>%
  ggplot(aes(x=male, y=female, label=antibiotic)) +
  geom_label() +
  geom_point() +
  geom_abline() +
  xlab('claims per 1,000 male beneficiaries') +
  ylab('claims per 1,000 female beneficiaries') +
  theme_minimal()
```
